---
title: "miscelaneus code"
author: "Nick Strayer"
date: "April 5, 2016"
output: pdf_document
---

Old code that I decided to leave out of the main explore markdown but we may desire again later. 

We have a large portion of lipids users who have no exposure. 
```{r}
ggplot(data, aes(avgexp)) + geom_histogram(binwidth = 0.05)
```

Now we check to see what the distribution of lipid use is. 

```{r}
ggplot(data, aes(x = factor(lipids))) + geom_bar(stat = "count") + labs(x = "Lipid Exposure")
```




So we can see more people did not recieve them than recieved them. 

```{r}
kable(summary(data[, y_n_cols]))
```


## Descriptive Statistics of Patients. 

First a quick function to generate a nice table output of `summary`. 
```{r}
#this is tedious
summaryTable <- function(variable){ kable(as.data.frame(as.list(summary(variable)))) }
```
### Ages
```{r}
summaryTable(data$age)
ggplot(data, aes(age)) + geom_histogram(binwidth = 2) + labs(x = "age")
```

---

### Gender
```{r}
summaryTable(data$gender)
ggplot(data, aes(x = factor(gender))) + geom_bar(stat = "count") + labs(x = "gender")
```

---


### Race
```{r}
#recode the ? race to unknown.
data[data$race == "?", "race"] = "UNKNOWN" 

summaryTable(data$race)
ggplot(data, aes(x = factor(race))) + geom_bar(stat = "count") + labs(x = "race")
```

---


### BMI
```{r}
summaryTable(data$bmi)
ggplot(data, aes(bmi)) + geom_histogram(binwidth = 3)
```

---


### Hospital Length of Stay
```{r}
summaryTable(data$hosp.los)
ggplot(data, aes(hosp.los)) + geom_histogram(binwidth = 3) + labs(x = "length of stay")
```

---


### Unit Length of Stay
```{r}
summaryTable(data$unit.los)
ggplot(data, aes(unit.los)) + geom_histogram(binwidth = 3) + labs(x = "length of stay")
```

---

### Glucose
```{r}
summaryTable(data$glucose)
ggplot(data, aes(glucose)) + geom_histogram(binwidth = 10) + labs(x = "glucose level")
```

---

### Hospital Deaths
```{r}
summaryTable(data$hosp.death)
ggplot(data, aes(x = factor(hosp.death))) + geom_bar(stat = "count") + labs(x = "Death")
```

A much greater amount of people didn't die. 

---


### Time on Ventilator
```{r}
summaryTable(data$ventdays.unit)
ggplot(data, aes(ventdays.unit)) + geom_histogram(binwidth = 3) + labs(x = "time on ventilator")
```

---

### Apache Score. 
```{r}
summaryTable(data$apache2)
ggplot(data, aes(apache2)) + geom_histogram(binwidth = 1) + labs(x = "APACHE score")
```
