---
title: "Lipids Exploration"
author: "Nick Strayer"
date: "April 4, 2016"
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
---
```{r, echo = FALSE, warning=FALSE,include=FALSE}
library(ggplot2)
library(knitr)
library(pastecs)
library(reshape)
```

First we read data in 
```{r}
data <- read.csv("/Users/Nick/Dropbox/vandy/regression/bios6312_final/data/ivfedata-2013_full.csv")
```

Now we recode the columns whos values are simply Ys and Ns to 1 and 0s respectively. 
```{r}
#y n columns
y_n_cols <- c("hosp.death","unit.death", "bsi.inf","eent.inf","gi.inf",
              "lri.inf","pneu.inf","ssi.inf","sst.inf","sys.inf","uti.inf")

#recode y and n values to 0=N, 1 = Y. 
data[, y_n_cols] = ifelse(data[, y_n_cols] == "N", 0, 1)
```

We have a large portion of lipids users who have no exposure. 
```{r}
ggplot(data, aes(avgexp)) + geom_histogram(binwidth = 0.05)
```

Because of this we will dichotomize on the exposure variable for logistic regression. 

```{r}
#make a binary column for if the patient recieved lipids during stay. 
data$lipids = ifelse(data$avgexp >  0.0, 1, 0)
y_n_cols = c("lipids",y_n_cols) #add this to the dichotomous variables now. 
```

---

\newpage

# Continious Variables

## Table of Descriptive Statistics
```{r}
continiuousVariables1 <- c("avgexp", "maxexp", "age", "bmi", "glucose", "hosp.los", "unit.los") 
continiuousVariables2 <- c("ventdays.hosp", "ventdays.unit", "ventfree.unit", "apache2") 

options(digits=2)
kable(stat.desc(data[, continiuousVariables1], basic=F))
kable(stat.desc(data[, continiuousVariables2], basic=F))
```

## Plot of histograms
```{r, warning=FALSE, fig.width=10, fig.height=9}
cvs <- c("avgexp", "maxexp", "age", "bmi", "glucose" , "hosp.los", "unit.los",
                          "ventdays.hosp", "ventdays.unit", "ventfree.unit", "apache2") 
cvplot <- data[, cvs]
cvplot <- melt(cvplot)

ggplot(cvplot, aes(value)) +
  geom_histogram(aes(y =..density..), col="black", alpha = 0.5) + 
  facet_wrap(~variable, scales = "free") + 
  geom_density(color = "steelblue")
```

---

\newpage

#Dichotomous Variables

## Outcomes plot: 
```{r, fig.width=7, fig.height=5.5}
tp <- melt(data[, y_n_cols])
ggplot(data.frame(table(tp)), aes(x = 1, y = Freq, fill = value)) + 
  geom_bar(stat="identity", position=position_stack()) + 
  facet_wrap( ~ variable) + labs(x = "")
```

```{r}
stats <- data.frame( no_occur = integer(), occur = integer())
for(i in 1:length(y_n_cols)) stats[i,] <- table(data[, y_n_cols[i]])
rownames(stats) <- c("Lipid Emulsion", "Death in Hospital","Death in Unit", "bsi.inf","eent.inf","gi.inf","lri.inf","pneu.inf","ssi.inf","sst.inf","sys.inf","uti.inf")
kable(stats, col.names= c("Occured", "Did not Occur"))
```


---

\newpage

## Big Scatterplots

### Looking at relationships between variables. 
```{r, fig.width=11, fig.height=10}
demoCols  <- c("avgexp",  "age",	"gender",	"bmi",	"hosp.los",	"hosp.death",	"ventdays.hosp",	"ventfree.unit",	"apache2")

plot(data[,demoCols])
```

Same thing but with the variables we chose as controls in our regressions. 
```{r, fig.width=10, fig.height=9}
controls = c("bmi", "glucose", "apache2", "race", "gender", "age") 
plot(data[, controls])
```
