---
title: "Lipids Exploration"
author: "Nick Strayer"
date: "April 4, 2016"
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
---
```{r, echo = FALSE, warning=FALSE}
library(ggplot2)
library(knitr)
library(pastecs)
```

First we read data in 
```{r}
data <- read.csv("/Users/Nick/Dropbox/vandy/regression/bios6312_final/data/ivfedata-2013_full.csv")
```

Now we recode the columns whos values are simply Ys and Ns to 1 and 0s respectively. 
```{r}
#y n columns
y_n_cols <- c("hosp.death","unit.death", "bsi.inf","eent.inf","gi.inf","lri.inf","pneu.inf","ssi.inf","sst.inf","sys.inf","uti.inf")

#recode y and n values to 0=N, 1 = Y. 
data[, y_n_cols] = ifelse(data[, y_n_cols] == "N", 0, 1)
```

We have a large portion of lipids users who have no exposure. 
```{r}
ggplot(data, aes(avgexp)) + geom_histogram(binwidth = 0.05)
```

Because of this we will dichotomize on the exposure variable for logistic regression. 

```{r}
#make a binary column for if the patient recieved lipids during stay. 
data$lipids = ifelse(data$avgexp >  0.0, 1, 0)
```

Now we check to see what the distribution of lipid use is. 

```{r}
ggplot(data, aes(x = factor(lipids))) + geom_bar(stat = "count") + labs(x = "Lipid Exposure")
```

So we can see more people did not recieve them than recieved them. 

```{r}
kable(summary(data[, y_n_cols]))
```

## Descriptive Statistics of Patients. 

First a quick function to generate a nice table output of `summary`. 
```{r}
#this is tedious
summaryTable <- function(variable){ kable(as.data.frame(as.list(summary(variable)))) }
```
### Ages
```{r}
summaryTable(data$age)
ggplot(data, aes(age)) + geom_histogram(binwidth = 2) + labs(x = "age")
```

---

### Gender
```{r}
summaryTable(data$gender)
ggplot(data, aes(x = factor(gender))) + geom_bar(stat = "count") + labs(x = "gender")
```

---


### Race
```{r}
#recode the ? race to unknown.
data[data$race == "?", "race"] = "UNKNOWN" 

summaryTable(data$race)
ggplot(data, aes(x = factor(race))) + geom_bar(stat = "count") + labs(x = "race")
```

---


### BMI
```{r}
summaryTable(data$bmi)
ggplot(data, aes(bmi)) + geom_histogram(binwidth = 3)
```

---


### Hospital Length of Stay
```{r}
summaryTable(data$hosp.los)
ggplot(data, aes(hosp.los)) + geom_histogram(binwidth = 3) + labs(x = "length of stay")
```

---


### Unit Length of Stay
```{r}
summaryTable(data$unit.los)
ggplot(data, aes(unit.los)) + geom_histogram(binwidth = 3) + labs(x = "length of stay")
```

---

### Glucose
```{r}
summaryTable(data$glucose)
ggplot(data, aes(glucose)) + geom_histogram(binwidth = 10) + labs(x = "glucose level")
```

---

### Hospital Deaths
```{r}
summaryTable(data$hosp.death)
ggplot(data, aes(x = factor(hosp.death))) + geom_bar(stat = "count") + labs(x = "Death")
```

A much greater amount of people didn't die. 

---


### Time on Ventilator
```{r}
summaryTable(data$ventdays.unit)
ggplot(data, aes(ventdays.unit)) + geom_histogram(binwidth = 3) + labs(x = "time on ventilator")
```

---


### Apache Score. 
```{r}
summaryTable(data$apache2)
ggplot(data, aes(apache2)) + geom_histogram(binwidth = 1) + labs(x = "APACHE score")
```

### Looking at relationships between variables. 
```{r, fig.width=10, fig.height=9}
demoCols  <- c("avgexp",	"age",	"gender",	"bmi",	"hosp.los",	"hosp.death",	"ventdays.hosp",	"ventfree.unit",	"apache2")

plot(data[,demoCols])
```

Same thing but with the variables we chose as controls in our regressions. 
```{r, fig.width=10, fig.height=9}
controls = c("bmi", "glucose", "apache2", "race", "gender", "age") 
plot(data[, controls])
```


## Outcomes plot: 
```{r, fig.width=7, fig.height=5.5}
library(reshape)
tp <- melt(data[, y_n_cols])
ggplot(data.frame(table(tp)), aes(x = 1, y = Freq, fill = value)) + 
  geom_bar(stat="identity", position=position_stack()) + 
  facet_wrap( ~ variable) + labs(x = "")
```

## Table of Descriptive Statistics
```{r}
continiuousVariables1 = c("avgexp", "maxexp", "age", "bmi", "hosp.los", "unit.los") 
continiuousVariables2 = c("ventdays.hosp", "ventdays.unit", "ventfree.unit", "apache2", "glucose" ) 

options(digits=2)
kable(stat.desc(data[, continiuousVariables1], basic=F))
kable(stat.desc(data[, continiuousVariables2], basic=F))
```

## Plot of histograms
```{r, warning=FALSE,fig.width=10, fig.height=9}
cvs= c("avgexp", "maxexp", "age", "bmi", "hosp.los", "unit.los",
                          "ventdays.hosp", "ventdays.unit", "ventfree.unit", "apache2", "glucose" ) 
cvplot <- data[, cvs]
cvplot <- melt(cvplot)

ggplot(cvplot, aes(value)) +
  geom_histogram(aes(y =..density..), col="black", alpha = 0.5) + 
  facet_wrap(~variable, scales = "free") + 
  geom_density(color = "steelblue")
```